### 2025-04-25 DGMM2025/computation.tex
Working on 3 (almost done)

As said in the introduction, any cell of $\mathcal{C}$ can be
characterized as a point with integer coordinates, aka \emph{lattice
point}. For instance, a standard way is to double the coordinates of
the centroid of the cell (often called its Khalimsky code). Now we
have to represent an arbitrary set $K$ of lattice points.

Let $(\ve_j)_{j=1,\dots}$ be the canonical basis of $\Z^d$.  We may
choose an arbitrary projection \emph{axis} $j \in
\{1,\ldots,d\}$. Letting $\pi_j$ be the projector along this axis, we
can collect all the cells of $K$ that projects onto the same
points. Let $p \in K$. Then $\pi^{-1}_j(p)$ is a set of cells having
all the same coordinates except along coordinate $j$. This set can be
ordered increasingly and stored as a list of integer intervals (see
Fig.~\ref{fig:lattice-representation} for an illustration).

\begin{figure}[t]
  \centering
  \input{lattice.tikz}
  \caption{\label{fig:lattice-representation} Representation of a
    2d cell complex (here the star of a given curve) as a lattice
    map. From 65 cells (with two coordinates) on the left, we get 11
    intervals on the right.}
\end{figure}

A \emph{sequence of intervals} is an ordered list of intervals $L =
([a_i,b_i])_{i=1,\ldots,n}$ such that $a_i \in \Z, b_i \in \Z, b_i + 1
< a_{i+1}$, i.e., disjoint intervals with at least a missing
integer. We denote by $\mathbb{L}$ the set of sequences of
intervals. An integer $k$ belongs to $L$ iff $\exists i$ such that
$a_i \Le k \Le b_i$. For a finite set of integers, there is a unique
sequence of intervals representing it.

A \emph{lattice map along axis $j$} is a set of pairs $(q,L_q)$, with
$q \in \Z^{d-1}$ and $L_q \in \mathbb{L}$. The lattice map $M_K$
\emph{represents} $K$ when, for any point $p \in K$, the pair
$(\pi_j(p), L_{\pi_j(p)})$ exists in $M_K$ and $p_j \in L_{\pi_j(p)}$,
with $p_j$ the $j$-th coordinate of $p$, and reciprocally, $\forall
(q,L_q) \in M_K, \forall i \in L_q, q+i\ve_j \in K$. A lattice map is
thus made of pairs $(q,L_q)$, where $q$ is the \emph{shift} while
$L_q$ is the \emph{intervals}. For a point $q$ which is a shift of
$M_K$, we write $M_K[q]$ for the corresponding intervals $L_q$.

The global algorithm for computing visibility is given in
Algorithm~\ref{alg:visibility}.  Given a digital input surface $C$ (a
subcomplex), we compute the lattice map $\Omega$ of $\Star{C}$ once,
generally along the axis where $C$ is the most elongated. In order to
compute visibility up to a given distance $r$, we consider all non
null primitive vectors $D$ with infinity norm no greater than $r$.
For each vector $\vv$, we compute the lattice map $M_\vv$ of its
intersected cells (as a vector from $\mathbf{0}$ to $\vv$ in
$\R^d$). The visibility along the direction $\vv$ reduces to computing
for each pair of $M_\vv$ the possible translations of intervals in
$\Omega$, and then intersecting all the translations.

%% \begin{definition}
%%   A sequence of intervals is an ordered list of intervals $L = ([a_i,b_i])_{i=1,\ldots,n}$ such that $b_i + 1 < a_{i+1}$, i.e., disjoint intervals with at least a missing integer.
%% \end{definition}

%% \begin{definition}
%%   A translation of an interval $A$ is defined as $A+t \coloneqq [a+t, b+t]$
%% \end{definition}

%% \begin{definition}
%%   A translation of a sequence of intervals $L$ is the sequence defined as $L+t \coloneqq \{L_1+t,\ldots,L_n+t\}$
%% \end{definition}

%% \begin{definition}
%%   All translations $T$ of an interval or a sequence of intervals $A$ in a sequence of intervals $L$ are defined as $ T \coloneqq \{ t, st. A+t \subset L\}$
%% \end{definition}

%% We will consider lattice maps as pairs of (shift, intervals), where shifts are points $p \in \R^{d-1}$. All coordinates are doubled, so that the represented cell $\forall c \in \R^d$ will have a dimension equal to $\sum_{i=1}^d \left(c[i]\mod d\right)$.
%% In figure~\ref{fig:lattice-representation}, we display a representation of lattice maps applied to a 2-d cell complex where all coordinates are already doubled. Note that the examples are in 2-d, the shown results are in 3-d and the algorithm is for n-d.

\begin{algorithm}
  \caption{Given a subcomplex $C$ and an integer $r$, returns the visibility from every point of $C$ up to distance $r$. The main axis is supposed to be $z$, while $x,y$ are the auxiliary axes.}
  \label{alg:visibility}
  \begin{algorithmic}
    \Function{Visibility}{$C$: Subcomplex, $r$: Integer}
    \State $\Omega \gets M_{\Call{Star}{C}}$ \Comment{Lattice map of the star of input subcomplex}
    \State $Directions \gets \Call{GetAllPrimalDirections}{r}$
    \State $V: \text{vector of boolean} \gets [0, \ldots, 0]$ \Comment{length $Size(Directions) \times \#C.pointels$}
    \State $low, high \gets \Call{BoundingBoxZ}{\Omega}$
    \ForAll{$\vv$ in $Directions$}
    \ForAll{shift $S$ in $\Omega$}
    \State $R \gets [low, high]$
    \ForAll{pair $P$ in $M_{\Call{Star}{[\mathbf{0},\mathbf{v}]}}$}
    \State $R \gets R \cap \Call{Translations}{P.intervals,\Omega[S + P.shift]}$
    \EndFor
    \State \Call{UpdateVisibility}{$V$, $R$}
    \EndFor
    \EndFor
    \State \Return $V$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{figure}
  \centering
  \input{algorithm.tikz}
  \caption{Evolution of the visibility check algorithm for a $(2,1)$
    vector. Green is the vector lattice map, black is the figure
    lattice map, red are the current intervals of positions where
    the visibility is still possible. The last red intervals are
    the visible positions. We travel the lattice maps from
    bottom-up and the found visibilities are drawn on the
    uppest figure.}
  \label{fig:visibility-algorithm-evolution}
\end{figure}

Figure~\ref{fig:visibility-algorithm-evolution} shows an example of
execution of this algorithm in an elementary case. It reduces to
determining all the possible inclusions by translation of sequences of
intervals in another sequence of intervals (note that for
pairwise visibility, it is enough to consider all the possible
inclusions by translation of one interval in another sequence of
intervals), and then intersecting progressively the results. Of
course, the process is stopped as soon as the intersection is
empty. Some results of visible points are displayed on
figure~\ref{fig:visibility-results} on 3d shapes.

%% In order to compute the visibility, we first compute $\Omega = \text{Star}(C)$. Then for each primal direction of
%% coordinates at most $r$, we look at all the possible positions this specific direction does link 2 visible points
%% of the cell complex. In order to do so, for each shift in $\Omega$, we compute the positions where points are
%% visible using the same method as presented in~\ref{fig:visibility-algorithm-evolution}. Some results of visible
%% points are present in figure~\ref{fig:visibility-results}.

\begin{algorithm}
  \caption{Given 2 lists of integer intervals $K$ and $L$, returns $K \cap L$}
  \label{alg:intersection}
  \begin{algorithmic}
    \Function{Intersection ($\cap$)}{\text{K, L}: Intervals}
    \State $R \gets \emptyset$; $k \gets 0$; $l \gets 0$;
    \While{$k < K.nbIntervals \And l < L.nbIntervals$}
    \State $[a,b] \gets K[k]$; $[c,d] \gets L[l]$
    \State $e \gets \max(a, c)$; $f \gets \min(b, d)$;
    \If{$e \leq f$}
    $R.append([e, f])$;
    \EndIf
    \If{$b \leq d$}
    $k \gets k+1$;
    \EndIf
    \If{$d \leq b$}
    $l \gets l+1$;
    \EndIf
    \EndWhile
    \State \Return $R$
    \EndFunction
  \end{algorithmic}
\end{algorithm}

To efficiently compute the intersection of two sequences of intervals
$K$ and $L$ (Algorithm~\ref{alg:intersection}), we go through the
first $2$ unvisited elements of the $2$ lists. If they do not overlap
(i.e.\ the start and end of the 1st interval are both smaller than the
start of the other), then we can discard the smaller interval. Else,
we construct the intersection of these 2 intervals as one of the
resulting intervals. We then can skip the interval with the smallest
end. If both intervals have the same end, then both intervals can be
skipped. Doing this computation until one of the lists is empty
returns the intersection of both lists of intervals.

% Insert visibility results images

\begin{figure}
  \centering
  \begin{tabular}{c c}
    \includegraphics[width=0.4\textwidth]{pictures/visibility_from_given_point_r_10} &
    \includegraphics[width=0.4\textwidth]{pictures/visibility_aware_of_features}
  \end{tabular}
  \caption{Examples of visibilities from a source point:
    (left) the visibility is stopped at the sharp edge, (right) the
    visibility does not cross the gap between the two parts.}
  \label{fig:visibility-results}
\end{figure}
%
%    \begin{figure}
%        \begin{center}
%            \includegraphics[width=0.8\textwidth]{pictures/visibility_from_given_point_r_10}
%            \caption{Visibility of a point on an edge of a fandisk}
%            \label{fig:visibility-fandisk}
%        \end{center}
%    \end{figure}
%    \begin{figure}
%        \begin{center}
%            \includegraphics[width=0.8\textwidth]{pictures/visibility_aware_of_features}
%            \caption{Visibility of a point on a torus knot (the visibility is feature-aware)}
%            \label{fig:visibility-torus-knot}
%        \end{center}
%    \end{figure}

% Quantitative results

We have plotted on Fig.\ref{fig:meanvisibility-computationComplexity}
the running times to compute global visibility of both \cite[Algorithm
  3]{lachaud:2022-jmiv} and our algorithm. Timings are very
similar. Our algorithm is faster for smaller maximal radii (up to 20),
while the other is slightly faster for bigger radii. We have also
measured the (discrete) average visibility distance along digitization
of smooth shapes (Fig.~\ref{fig:meanvisibility-gridstep}). It follows
a law proportionnal to $\sqrt{h}$. As shown in the next section, for
real 3d images, a maximal radius below 10 is sufficient for normal and
curvature estimation.

\begin{figure}[t]
  \centering
  \input{pointelscomptime.tikz}
  \caption{ \label{fig:meanvisibility-computationComplexity}Computation
    time of visibility as a function of the number of pointels. We
    compare the running time of the naive breadth first visibility
    algorithm against our algorithm using intervals. We input the same
    maximal radius to both versions (10, 20, 30) and test it on 6
    different figures (goursat, torus, rcube, sphere9, leopold,
    ``D20'') with various gridsteps, ($\#$ pointels ranging from 520
    to 390235.}
\end{figure}

\begin{figure}
  \centering
  \input{mean-distance.tikz}
  \caption{\label{fig:meanvisibility-gridstep} Mean distance of visibility as a function of the gridstep $h$ (see text).}
\end{figure}